Asymmetric relationships and their effects on coexistence

### 2023-08-01 paper/Paper.Rnw
mostly trimming to get under the 5000 word length for the main text

%more stuff that applies to both
For both versions of the lottery model, 
fecundities were taken to directly reflect 
the environment, i.e., the $E_i$ of Theory was set equal to $B_i$. 
Competition, $C_i(t)$, in the lottery model occurs at the stage of juveniles occupying open sites:
$C_i$ was taken to be the number of new offspring
divided by the number of open sites, $C_i(t)=( B_1(t)N_1(t)+B_2(t)N_2(t) )/(\delta N$), which
does not depend on $i$. These choices for $E_i(t)$ and $C_i(t)$ fulfill the assumption of Theory 
that $r_i(t)$ is an increasing function of $E_i(t)$ and a decreasing function of $C_i(t)$ 
(SI section \ref{SI-SIsect:TheoryForLottery}).

@@

%Plotting shortcuts for the log-normal fecundities model
To explore, using the log-normal fecundities lottery model, the relative magnitudes of the contributions 
to coexistence that come from ATAs versus other mechanisms, we plotted the terms of the decomposition 
(\ref{eq:full_decomp}) for that model against the model parameters $\mu_1$, $\mu_2$, $\sigma$ and $\delta$.
We labeled regions of the plot which corresponded to ATA rescue or exclusion, i.e., regions for which
GWR and GWR-$\Delta_i^{[EC]}$ had opposite signs.
Some mathematical shortcuts, summarized here, simplified plotting. 
First, we assumed, without loss of generality, that $\mu_1 \leq \mu_2$. To understand coexistence, 
it was then sufficient to decompose GWR of the weaker competitor, species 1. 
Second, we showed that, for the log-normal fecundities
lottery model, none of the components of (\ref{eq:full_decomp}) depends independently on $\mu_1$ and $\mu_2$;
instead they depend only on $\mu_1-\mu_2$ (SI section \ref{SI-SIsect:TheoryForLottery}). 
So we plotted against $\mu_1 - \mu_2$. 
Third, we showed that the values of all components in (\ref{eq:full_decomp}) were the same 
for both our left- and right-tail associated noises (SI section \ref{SI-SIsect:TheoryForLottery}). 
So we only generated plots 
for left-tail associated noise.

For each combination of $\delta=0.2, 0.4, 0.6$ and $\mu_1-\mu_2=0,-0.5,-2,-4$,
we plotted the components of (\ref{eq:full_decomp}) against $\sigma$ for $\sigma$ ranging 
from $0$ to $7$. \cite{Chesson_1981} showed that coexistence is impossible for $\delta=1$, and it
is known that the storage effect (and therefore its ATA component) is weak for larger $\delta$, so we 
only considered values of $\delta$ up to $0.6$. Some of 
the $\mu$ differences and $\sigma$ values we considered were very unrealistic (see Results),
so we considered after plotting whether conclusions about the importance of ATAs depended 
on realistic values of parameters.
The simulation methods of \cite{Ellner_2019} were used to compute the components of (\ref{eq:full_decomp}).
Computationally efficient means of performing those simulations are in 
SI section \ref{SI-sect:efficient_computation}.

%Then a similar para for the beta-fecundity lottery model.
To explore, using the beta fecundities lottery model, the relative magnitudes of the contributions 
to coexistence that come from ATAs versus other mechanisms, we plotted the terms of the decomposition 
(\ref{eq:full_decomp}) for that model against the model parameters $\eta_1$, $\eta_2$, and $\delta$,
again labelling regions corresponding to ATA rescue or exclusion.
Some mathematical shortcuts, summarized here, simplified plotting. 
First, we assumed, without loss of generality, that $\eta_1 \leq \eta_2$. To understand coexistence, 
it was then sufficient to decompose GWR of the weaker competitor, species 1. 
Second, we showed that, for the beta fecundities
lottery model, none of the components of (\ref{eq:full_decomp}) depends independently on $\eta_1$ and $\eta_2$;
instead they depend only on $\eta_2/\eta_1$ (SI section \ref{SI-SIsect:TheoryForLottery}). 
So we plotted against $\eta_2/\eta_1$.

For each combination of $\delta=0.2, 0.4, 0.6$
and for left- and right-tail associated noise, we plotted 
the components of (\ref{eq:full_decomp}) against the upper bound ratio $\eta_2/\eta_1$, using 
the fairly realistic (see Results) range $1 \leq \eta_2/\eta_1 \leq 5$. 
The simulation methods of \cite{Ellner_2019} were again used.
Computationally efficient means of performing these simulations are in 
SI section \ref{SI-sect:efficient_computation}.

%Trait different space stuff. I don't think I need it anymore but keeping around until I finish writing the Discussion,
%in case some of it turns out to be useful there.
% To achieve goal G3 from the Introduction, i.e., to explore, using the lottery model, the likelihood that species can coexist with
% ATAs but not without, or vice versa, we plotted GWR, $\olr_{i \bs i}=\olr_{i \bs i} -q_{ij}\olr_{j \bs i}$, 
% as well as
% GWR minus its ATA contribution, $\olr_{i \bs i} -\Delta_i^{[EC]}$, again for species $i=1$, the weaker competitor. 
% For fixed values of $\sigma$ and $\delta$ and for left-tail associated noise, 
% we were interested in the values of $\mu_1-\mu_2$ for which one of these
% quantities is positive and the other is negative; this set of values characterizes the species which can coexist
% with ATAs, but not without, or vice versa. We define the ``trait-difference space'' of values $\mu_1-\mu_2$. 
% Points of the trait-difference space characterize possible combinations of traits of the two 
% competing species in our model. For fixed values of 
% $\delta$ and $\sigma$, we defined the subset, $\psi_{\text{ATA}}(\delta,\sigma)$, of 
% trait-difference space to be the
% values $\mu_1 - \mu_2$ for which GWR, $\olr_{i \bs i}$, is positive. This 
% is the subspace for which the two lottery model species coexist in the presence of ATAs.
% We also defined the subset, $\psi_{\lnot\text{ATA}}(\delta,\sigma)$, of trait-difference space to be the values
% $\mu_1 - \mu_2$ for which the metric $\olr_{i \bs i} -\Delta_i^{[EC]}$ is positive.
% This is the subspace for which the two lottery model species coexist in the absence of 
% ATAs. The symbol $\lnot$ denotes logical negation, so $\lnot \text{ATA}$ indicates the
% absence of ATAs. We defined the ``ATA rescue space" associated with given values of $\delta$ and $\sigma$ to be the 
% set difference $\psi_{\text{ATA}}(\delta,\sigma) \setminus \psi_{\lnot\text{ATA}}(\delta,\sigma)$, i.e., the
% values $\mu_1 - \mu_2$ for which species coexist in the presence of ATAs, but not in their absence. 
% We define ``ATA exclusion space'' to be the set difference 
% $\psi_{\lnot\text{ATA}}(\delta,\sigma) \setminus \psi_{\text{ATA}}(\delta,\sigma)$,
% i.e., $\mu_1 - \mu_2$ values for which coexistence occurs in the absence, but not the presence, of ATAs.
% On plots of $\olr_{i \bs i}$ and $\olr_{i \bs i} -\Delta_i^{[EC]}$
% against $\mu_1 - \mu_2$, we highlighted the ATA rescue and exclusion spaces. We plotted against
% $\mu_1 - \mu_2$ for all combinations of $\delta=0.2, 0.4, 0.6, 0.8$ and $\sigma=2,4,5,6$.

@@

To achieve goal G3 from the Introduction, i.e., to provide an empirical example of species 
coexistence which highlights the role ATAs can play, we used a laboratory chemostat
system of freshwater diatoms explored by \cite{Descamps_2005}.
Those authors used measurements of growth rates of \emph{Fragilaria crotonensis} and 
\emph{Cyclotella pseudostelligera} over a range of temperatures to parametrize a resource
competition model. The model is 
\begin{align}
\frac{dS}{dt} &= D(S_0 - S) - Q_1x_1 \frac{V_1 S}{K_1 + S} - Q_2 x_2 \frac{V_2 S}{K_2 + S} \\
\frac{dx_j}{dt} &= x_j \frac{V_j S}{K_j+S}-D x_j, 
\end{align}
for $j=1,2$, where $S$ is an extracellular silicate concentration in the chemostat, $x_1$ is the population density
of \emph{F. crotonensis} and $x_2$ is that of \emph{C. pseudostelligera}, 
$D$ is the chemostat outflow rate, $S_0$ is concentration of silicates in the chemostat
inflow, the $Q_j$ are amounts of silicate per cell, the $V_j$ are maximum population growth rates, and
the $K_j$ are half-saturation constants for the dependence of growth rates on nutrient concentration.
The temperature-dependent parameters $Q_j$, $V_j$, and $K_j$ were obtained by 
\cite{Descamps_2005} through batch experiments and curve fitting or interpolation.
\cite{Descamps_2005} showed that permanent coexistence of the two species occurs when 
temperature fluctuates periodically, 
$\theta(t)=\theta_0 +  a\sin(2\pi t/P)$, 
with parameters $\theta_0 = 18^\circ$C, $a = 6^\circ$C, $P = 60$d. 
\cite{Ellner_2016, Ellner_2019} broke down coexistence into contributing mechanisms
according to (\ref{eq:previous_decomp}) for these same values of
$\theta_0$, $a$ and $P$. We further decompose
coexistence according to (\ref{eq:full_decomp}), and we do so for ranges of values
of $\theta_0$, $a$ and $P$. In contrast to the lottery model case, for which simplifying assumptions
made it sufficient to decompose only the GWR of species 1, for this
example we considered $\olr_{i \bs i} -q_{ij}\olr_{j \bs i}$ for both $i=1$, $j=2$ and for
$i=2$, $j=1$. Further details of the model setup are in SI section \ref{SI-sect:diatom_model}
and specifics of how our decomposition extends to this empirical case are in 
SI section \ref{SI-sect:Empircal_app}.

@@

%One-parameters plots
To display results for goal G3, we started by plotting the terms of 
(\ref{eq:full_decomp}) against one of the parameters 
$\theta_0$, $a$ and $P$ at a time, while keeping the other two fixed at the values
used by \cite{Descamps_2005}. We again labelled regions of parameter space corresponding to
ATA rescue or exclusion.

%Two-parameter plots
We also generated two-dimensional plots, where
two of the parameters $\theta_0$, $a$ and $P$ were varied and the third was kept fixed
at the value used by \cite{Descamps_2005}. For the two-dimensional plots, the quantity  
$\Delta_i^{[EC]}/|\text{GWR}|$ was displayed using color, to convey the 
importance of the ATA contribution relative to the whole GWR. Larger magnitudes
of this quantity indicate greater importance of ATAs. The case
$\Delta_i^{[EC]}>\text{GWR}>0$ (which produces $\Delta_i^{[EC]}/|\text{GWR}|>1$)
corresponds to ATA rescue, and the case $\Delta_i^{[EC]}<\text{GWR}<0$
(which produces $\Delta_i^{[EC]}/|\text{GWR}| < -1$) corresponds to ATA exclusion.
Thus the value of $\Delta_i^{[EC]}/|\text{GWR}|$ indicates the importance of ATAs
and also whether ATA rescue or exclusion occurred.

@@

To begin fulfilling goal G2 of the Introduction (to explore, using the lottery model, the relative magnitudes 
of the contributions to coexistence that come from ATAs compared to other mechanisms; and to see if ATA rescue 
or ATA exclusion can occur), Fig.
\ref{fig:fig3} shows that, for the log-normal fecundities lottery model, 
the ATA term of (\ref{eq:full_decomp}) sometimes, but not always, has 
magnitude comparable to the magnitudes of other terms of the extended MCT decomposition. 
Magnitudes of $\Delta_i^{[EC]}$ tended to be larger, relative to the magnitudes of the 
other terms of (\ref{eq:full_decomp}), for smaller-magnitude differences between $\mu_1$ and 
$\mu_2$, i.e., for closer-to-equal competitors. 
In the $\mu_1=\mu_2$ case (panels a-c), for which the two species were equal 
competitors, storage effects ($\Delta_i^{[EC]}+\Delta_i^{[E\nat C]}$) 
were the only means by which coexistence could occur, and ATA effects
tended to be negative, weakening total storage effects.

The ATA term 
was also often, or always (for the parameters we considered), comparable in magnitude to the other terms of 
the decomposition (\ref{eq:full_decomp}) for the beta fecundities lottery model 
(Fig. \ref{fig:fig4}). The term contributed negatively to coexistence for environmental noise/fecundities
showing left-tailed association, and contributed positively for noise/fecundities showing right-tailed 
association.

Our results also show that ATAs can make the difference between coexistence 
and failure of two species to coexist: both ``ATA rescue'' and ``ATA
exclusion'' (see Theory) are possible. Starting with the log-normal fecundities 
lottery model, for some of the panels on Fig. \ref{fig:fig3}, the GWR line falls between the ATA
line and the $x$ axis for some values of $\sigma$, e.g., see panel i.
Such cases correspond to parameter combinations
for which the presence or absence of ATAs determine whether the species coexist or not. In the 
case of panel i, for values of $\sigma$ around $6$, 
species 1 is excluded with ATAs but the two species can coexist without ATAs. 
Fig. \ref{fig:fig4} highlights additional examples using the beta fecundities lottery model; 
examples were more common than for the log-normal fecundities model.

For the log-normal fecundities lottery model, the ATA term from (\ref{eq:full_decomp}) tended
only to be comparable in magnitude to the other terms, and ATA exclusion or rescue tended only to
occur, for unrealistic parameters values; but ATA terms were large and ATA rescue and exclusion occurred
for many realistic parameter values for the beta fecundities lottery model. For instance, the ATA term in 
Fig. \ref{fig:fig3} had fairly 
small magnitude, compared to the other terms of (\ref{eq:full_decomp}), for values of $\sigma$ less than 
about $3$. And ATA rescue and exclusion only occurred for $\sigma$ greater than about $3$.
The standard deviation of fecundity for species 1 when 
$\mu_1=0$ and $\sigma=3$ was greater than $8000$, so values of $\sigma$ greater than $3$ are 
probably unrealistic. Thus ATAs seem to be unimportant for coexistence
for the log-normal fecundities lottery model with realistic parameter values. However, 
all the $\eta_2/\eta_1$ ratio values we plotted were fairly realistic, corresponding to
upper-bound fecundities that were up to $5$ times higher for species 2 than for species 1. And 
the ATA term plotted in Fig. \ref{fig:fig4} was comparable in magnitude 
to the other components of (\ref{eq:full_decomp}) across the whole range of parameters
considered. And both ATA rescue and exclusion occurred for values of $\eta_2/\eta_1$ between $1$ and $2$.
Apparently the question of whether ATAs are important for realistic parameter 
values depends on specifics.

@@

Starting to fulfill goal G3 of the Introduction (to provide an empirical example of species coexistence 
which highlights the role ATAs can play), Fig. \ref{fig:fig5}a-c 
show that ATA contributions to coexistence often have magnitudes comparable to the 
magnitudes of other terms of the extended MCT decomposition, and that 
ATA rescue occurs for some combinations of parameters for this system.

Plotting against two parameters at a time, Fig. \ref{fig:fig5}d-f reinforce the 
same conclusions. The plotted quantity $\Delta_i^{[EC]}/|\text{GWR}|$ was often large,
indicating that ATAs were often a substantial contributor to whether the diatom species 
coexist. For instance, well over half the area of Fig. \ref{fig:fig5}f showed values of 
$\Delta_i^{[EC]}/|\text{GWR}|$ bigger than $1/4$, with large portions of
Fig. \ref{fig:fig5}d-e satisfying the same criterion. Substantial portions of 
panels d-f also show values of $\Delta_i^{[EC]}/|\text{GWR}|$ bigger than $1$,
corresponding to ATA rescue. ATA exclusion did not occur for this system for the 
parameter ranges we considered.

@@

%Summary-type paragraph of main results
We produced and applied a new extension of modern coexistence theory (MCT) that 
quantifies the impact of asymmetric tail associations (ATAs) on species coexistence. 
Building upon a recent, simulation-based approach to MCT \citep{Ellner_2016, Ellner_2019},
we decomposed storage effects into two new mechanisms, the contribution of ATAs to coexistence,
and the contribution of $EC$ correlation \emph{per se}.
Applications of our extended theory to the lottery model and to a laboratory diatom system demonstrated that  
ATA contributions to storage effects and to species coexistence 
can often be comparable in magnitude to the contributions of previously recognized mechanisms.
Our results add
coexistence and biodiversity maintenance to the list of ecological theories and phenomena on which 
ATAs have been shown to have an important impact \citep{Ghosh_2020_AER, Ghosh_2020_Eco, Ghosh_2020_ecolevol, Ghosh_2021}.  
Though storage effects are typically described as the contribution
of $EC$ \emph{covariance} to coexistence \citep{Chesson_2012, Ellner_2016, Ellner_2019},
our results suggest that storage effects should be redefined to constitute the 
contribution of \emph{associations} between $E$ and $C$, generally, including both ATA effects and
correlation \emph{per se}. The prior terminology, \emph{covariance}, is typically construed 
as reflecting the linear relationship between variables, ignoring complexities of the association,
such as ATAs, which we now know can be important. ATAs contributions to coexistence were less important than other mechanisms 
in some contexts (the log-normal fecundities lottery model), but were very important in other contexts
(the beta fecundities lottery model), including our empirically
based example (the diatom system). 
Future work should seek to understand precisely when ATAs are or are not important
for coexistence. Our work demonstrates the potential for ATAs to be an important mechanism of coexistence, generally.

%Para continuing to try to build intuition about how ATAs work
We return to the task, begun in the Introduction, of building intuition for how
ATA effects work. 
Just as for storage effects [\cite{Chesson_1994}, Box 2 of \cite{StumpVasseur2023}], 
temporal niche differentiation is necessary for ATA
effects (SI section \ref{SI-SIsect:temporal_niche_diff}). 
If temporal niche differentiation occurs
but is partial, as will be typical, $E_i$ and $C_{i \bs i}$ for a 
species $i$ will be positively but imperfectly associated, 
and the association may be asymmetric. 
Storage events occur for species $i$ when the positive effects of $E_i$ overwhelm the negative 
effects of $C_{i \bs i}$, producing large values of $r_i(E_i,C_{i \bs i})$
which are not completely offset by population declines at other times. 
For instance, for the lottery model,
$r_i(E_i,C_{i \bs i})=\log(1-\delta+E_i/C_{i \bs i})$ can get large whenever
$C_{i \bs i}$ is a small fraction of $E_i$, but can never get smaller
than $\log(1-\delta)$. 
The potential for storage events can differ when $E_i$ and $C_{i \bs i}$ are both
above their median values compared to when they are both below median. For instance,
the ratio $E_i/C_{i \bs i}$, and therefore the growth rate 
$\log(1-\delta+E_i/C_{i \bs i})$, can both get very big, corresponding to
rapid growth, if $C_{i \bs i}$ is close to $0$, even if $E_i$ is fairly small.
In contrast, if $E_i$ and $C_{i \bs i}$ are both in the right portions of their distributions, 
so that $C_{i \bs i}$ is bounded away from $0$, there may be much sharper limits on the 
ratio $E_i/C_{i \bs i}$ and the corresponding growth rate.
ATA effects quantify the extent to which asymmetries of association between $E$ and 
$C$ are aligned with varying potential for storage across the $E,C$ distribution. 
ATA effects help quantify different degrees of temporal niche differentiation 
occurring under different sets of environmental conditions, possibly via
different mechanisms.
Further details and examples of this intuition 
are in SI section \ref{SI-SIsect:intuition_examples} and SI 
Figs \ref{SI-fig:beta_fecundidities_explanatory}-\ref{SI-fig:lognormal_fecundidities_explanatory}, 
where comparisons are also made between our two lottery models.

>>>>

%more stuff that applies to both
For both versions of the lottery model, 
fecundities directly reflected 
the environment, i.e., the $E_i$ of Theory was set equal to $B_i$. 
Competition, $C_i(t)$, in the lottery model occurs at the stage of juveniles occupying open sites:
$C_i$ was taken to be the number of new offspring
divided by the number of open sites, $C_i(t)=( B_1(t)N_1(t)+B_2(t)N_2(t) )/(\delta N$), which
does not depend on $i$. These choices for $E_i(t)$ and $C_i(t)$ fulfill the assumption of Theory 
that $r_i(t)$ is an increasing function of $E_i(t)$ and a decreasing function of $C_i(t)$ 
(SI section \ref{SI-SIsect:TheoryForLottery}).

@@

%Plotting shortcuts for the log-normal fecundities model
To explore, using the log-normal fecundities lottery model, the relative magnitudes of the contributions 
to coexistence of ATAs versus other mechanisms, we plotted the terms of the decomposition 
(\ref{eq:full_decomp}) against model parameters $\mu_1$, $\mu_2$, $\sigma$ and $\delta$.
We labeled regions of the plot which corresponded to ATA rescue or exclusion, i.e., regions for which
GWR and GWR-$\Delta_i^{[EC]}$ had opposite signs.
Some mathematical shortcuts, summarized here, simplified plotting. 
First, we assumed, without loss of generality, that $\mu_1 \leq \mu_2$. To understand coexistence, 
it was then sufficient to decompose GWR of the weaker competitor, species 1. 
Second, we showed that none of the components of (\ref{eq:full_decomp}) depends independently on $\mu_1$ and $\mu_2$;
instead they depend only on $\mu_1-\mu_2$ (SI section \ref{SI-SIsect:TheoryForLottery}). 
So we plotted against $\mu_1 - \mu_2$. 
Third, we showed that the values of all components in (\ref{eq:full_decomp}) were the same 
for both our left- and right-tail associated noises (SI section \ref{SI-SIsect:TheoryForLottery}). 
So we only generated plots for left-tail associated noise.

For each combination of $\delta=0.2, 0.4, 0.6$ and $\mu_1-\mu_2=0,-0.5,-2,-4$,
we plotted the components of (\ref{eq:full_decomp}) against $\sigma$ for $\sigma$ 
from $0$ to $7$. \cite{Chesson_1981} showed that coexistence is impossible for $\delta=1$, and it
is known that the storage effect is weak for larger $\delta$, so we 
only considered $\delta$ up to $0.6$. Some of 
the $\mu$ differences and $\sigma$ values we considered were unrealistic (see Results),
so we considered after plotting whether conclusions depended 
on realistic values of parameters.
Simulation methods of \cite{Ellner_2019} were used to compute the components of (\ref{eq:full_decomp}).
Computationally efficient means of performing those simulations are in 
SI section \ref{SI-sect:efficient_computation}.

%Then a similar para for the beta-fecundity lottery model.
To explore, using the beta fecundities lottery model, the relative magnitudes of the contributions 
to coexistence of ATAs versus other mechanisms, we plotted the terms of the decomposition 
(\ref{eq:full_decomp}) against model parameters $\eta_1$, $\eta_2$, and $\delta$,
again labelling regions of ATA rescue or exclusion.
Mathematical shortcuts, summarized here, simplified plotting. 
First, we assumed, without loss of generality, that $\eta_1 \leq \eta_2$. 
It was then sufficient to decompose GWR of the weaker competitor, species 1. 
Second, we showed that none of the components of (\ref{eq:full_decomp}) depends independently on $\eta_1$ and $\eta_2$;
instead they depend only on $\eta_2/\eta_1$ (SI section \ref{SI-SIsect:TheoryForLottery}). 
So we plotted against $\eta_2/\eta_1$.

For each combination of $\delta=0.2, 0.4, 0.6$
and for left- and right-tail associated noise, we plotted 
the components of (\ref{eq:full_decomp}) against the upper bound ratio $\eta_2/\eta_1$, using 
the fairly realistic (see Results) range $1 \leq \eta_2/\eta_1 \leq 5$. 
Simulation methods of \cite{Ellner_2019} were again used, with 
computationally efficient means of simulating in 
SI section \ref{SI-sect:efficient_computation}.

@@

To achieve goal G3 from the Introduction, i.e., to provide an empirical example of species 
coexistence which highlights the role ATAs can play, we used a laboratory chemostat
system of freshwater diatoms explored by \cite{Descamps_2005}.
Those authors used measurements of growth rates of \emph{Fragilaria crotonensis} and 
\emph{Cyclotella pseudostelligera} over a range of temperatures to parameterize a resource
competition model. The model is 
\begin{align}
\frac{dS}{dt} &= D(S_0 - S) - Q_1x_1 \frac{V_1 S}{K_1 + S} - Q_2 x_2 \frac{V_2 S}{K_2 + S} \\
\frac{dx_j}{dt} &= x_j \frac{V_j S}{K_j+S}-D x_j, 
\end{align}
for $j=1,2$, where $S$ is an extracellular silicate concentration in the chemostat, $x_1$ is the population density
of \emph{F. crotonensis} and $x_2$ is that of \emph{C. pseudostelligera}, 
$D$ is the chemostat outflow rate, $S_0$ is concentration of silicates in the chemostat
inflow, the $Q_j$ are amounts of silicate per cell, the $V_j$ are maximum population growth rates, and
the $K_j$ are half-saturation constants for the dependence of growth rates on nutrient concentration.
The temperature-dependent parameters $Q_j$, $V_j$, and $K_j$ were obtained by 
\cite{Descamps_2005} through batch experiments and curve fitting or interpolation.
\cite{Descamps_2005} showed that permanent coexistence of the two species occurs when 
temperature fluctuates periodically, 
$\theta(t)=\theta_0 +  a\sin(2\pi t/P)$, 
with parameters $\theta_0 = 18^\circ$C, $a = 6^\circ$C, $P = 60$d. 
\cite{Ellner_2016, Ellner_2019} broke down coexistence into contributing mechanisms
according to (\ref{eq:previous_decomp}) for these same values of
$\theta_0$, $a$ and $P$. We further decompose
coexistence according to (\ref{eq:full_decomp}), and we do so for ranges of values
of $\theta_0$, $a$ and $P$. In contrast to the lottery model case, for which simplifying assumptions
made it sufficient to decompose only the GWR of species 1, for this
example we considered $\olr_{i \bs i} -q_{ij}\olr_{j \bs i}$ for both $i=1$, $j=2$ and for
$i=2$, $j=1$. Details of the model setup are in SI section \ref{SI-sect:diatom_model}
and specifics of how our decomposition was computed are in 
SI section \ref{SI-sect:Empircal_app}.

@@

%One-parameters plots
To display results for G3, we started by plotting the terms of 
(\ref{eq:full_decomp}) against one of the parameters 
$\theta_0$, $a$ and $P$ at a time, keeping the other two fixed at the values
used by \cite{Descamps_2005}. We again labelled regions of parameter space corresponding to
ATA rescue or exclusion.

%Two-parameter plots
We also generated two-dimensional plots, where
two of the parameters $\theta_0$, $a$ and $P$ were varied and the third was kept fixed. 
For those plots, the quantity  
$\Delta_i^{[EC]}/|\text{GWR}|$ was displayed using color. Larger magnitudes
of this quantity indicate greater importance of ATAs. The case
$\Delta_i^{[EC]}>\text{GWR}>0$ (which produces $\Delta_i^{[EC]}/|\text{GWR}|>1$)
corresponds to ATA rescue, and $\Delta_i^{[EC]}<\text{GWR}<0$
(which produces $\Delta_i^{[EC]}/|\text{GWR}| < -1$) corresponds to ATA exclusion.
Thus $\Delta_i^{[EC]}/|\text{GWR}|$ indicates the importance of ATAs
and also whether ATA rescue or exclusion occurred.

@@

To begin fulfilling goal G2 of the Introduction (to explore, using the lottery model, the relative magnitudes 
of the contributions of ATAs compared to other coexistence mechanisms; and to see if ATA rescue 
or exclusion occur), Fig.
\ref{fig:fig3} shows that, for the log-normal fecundities lottery model, 
the ATA term of (\ref{eq:full_decomp}) sometimes, but not always, has 
magnitude comparable to the magnitudes of other terms. 
Magnitudes of $\Delta_i^{[EC]}$ tended to be larger, relative to the magnitudes of the 
other terms of (\ref{eq:full_decomp}), for smaller-magnitude differences between $\mu_1$ and 
$\mu_2$, i.e., for closer-to-equal competitors. 
In the $\mu_1=\mu_2$ case (panels a-c), for which the two species were equal 
competitors, storage effects ($\Delta_i^{[EC]}+\Delta_i^{[E\nat C]}$) 
were the only means by which coexistence could occur, and ATA effects
tended to be negative, weakening storage effects.

For the beta fecundities lottery model, the ATA term 
was often or always, for the parameters we considered, comparable in magnitude to the other terms of 
the decomposition (\ref{eq:full_decomp}) (Fig. \ref{fig:fig4}). The term contributed negatively to 
coexistence for environmental noise/fecundities
showing left-tailed association, and positively for right-tailed 
association.

Our results also show that ATAs can make the difference between coexistence 
and competitive exclusion: both ``ATA rescue'' and ``ATA
exclusion'' (see Theory) are possible. Starting with the log-normal fecundities 
lottery model, for some of the panels on Fig. \ref{fig:fig3}, the GWR line falls between the ATA
line and the $x$ axis for some values of $\sigma$, e.g., panel i.
For such cases, the presence or absence of ATAs determine whether the species coexist or not. 
For panel i, for values of $\sigma$ around $6$, 
species 1 is excluded with ATAs but the two species coexist without ATAs. 
Fig. \ref{fig:fig4} highlights additional examples using the beta fecundities lottery model,
where examples were more common than for the log-normal fecundities model.

Contrasting the two models, ATAs were only important for the log-normal fecundities lottery model 
when that model was parameterized unrealistically; but ATAs were important for a wide range of 
realistic parameter values for the beta fecundities model.
For instance, the ATA term in Fig. \ref{fig:fig3} had large magnitude compared to the other
terms of (\ref{eq:full_decomp}), and ATA rescue and exclusion occurred, only for 
$\sigma$ greater than about $3$. 
The standard deviation of fecundity for species 1 when 
$\mu_1=0$ and $\sigma=3$ was greater than $8000$, so values of $\sigma$ greater than $3$ are 
probably unrealistic. 
However, all the $\eta_2/\eta_1$ ratio values we plotted for the beta fecundities model 
were fairly realistic, corresponding to
upper-bound fecundities that were up to $5$ times higher for species 2 than for species 1.  
The ATA term in Fig. \ref{fig:fig4} was comparable in magnitude 
to other components of (\ref{eq:full_decomp}) across the whole range of parameters
considered. And ATA rescue and exclusion occurred for values of $\eta_2/\eta_1$ between $1$ and $2$.
Apparently the question of whether ATAs are important for realistic parameter 
values depends on specifics.

@@

Starting to fulfill goal G3 of the Introduction (to provide an empirical example of species coexistence 
which highlights the role ATAs can play), Fig. \ref{fig:fig5}a-c 
show that ATA contributions to coexistence often have magnitudes comparable to those 
of other terms of the extended MCT decomposition, and that 
ATA rescue occurs for some parameters for this system.

Plotting against two parameters at a time, Fig. \ref{fig:fig5}d-f reinforce the 
same conclusions. The quantity $\Delta_i^{[EC]}/|\text{GWR}|$ was often large,
indicating that ATAs were a substantial contributor to coexistence. 
For instance, well over half the area of Fig. \ref{fig:fig5}f showed values of 
$\Delta_i^{[EC]}/|\text{GWR}|$ bigger than $1/4$, with large portions of
Fig. \ref{fig:fig5}d-e satisfying the same criterion. Substantial portions of 
panels d-f also show values of $\Delta_i^{[EC]}/|\text{GWR}|$ bigger than $1$,
corresponding to ATA rescue. ATA exclusion did not occur for this system for the 
parameters we considered.

@@

%Summary-type paragraph of main results
We developed and applied a new extension of modern coexistence theory (MCT) that 
quantifies the impact of asymmetric tail associations (ATAs) on species coexistence. 
Building upon a recent, simulation-based approach to MCT \citep{Ellner_2016, Ellner_2019},
we decomposed storage effects into two new mechanisms, the contribution of ATAs to coexistence,
and the contribution of $EC$ correlation \emph{per se}.
Applications of our extended theory to the lottery model and to a laboratory diatom system demonstrated that  
ATA contributions to storage effects and to species coexistence 
can be comparable in magnitude to contributions of previously recognized mechanisms.
Our results add
coexistence to the list of ecological phenomena on which 
ATAs have been shown to have an important impact \citep{Ghosh_2020_AER, Ghosh_2020_Eco, Ghosh_2020_ecolevol, Ghosh_2021}.  
Though storage effects are typically described as the contribution
of $EC$ \emph{covariance} to coexistence \citep{Chesson_2012, Ellner_2016, Ellner_2019},
our results suggest that storage effects should be redefined as the 
contribution of \emph{associations} between $E$ and $C$, generally, including ATA effects and
correlation \emph{per se}. The prior terminology, \emph{covariance}, is typically construed 
as reflecting the linear relationship between variables, ignoring complexities of the association
such as ATAs, which we now know can be important. ATAs contributions to coexistence were less important than other mechanisms 
in some contexts, but were very important in other contexts, including our empirically
based example (the diatom system).
Our work therefore demonstrates the potential for ATAs to be an important mechanism of coexistence, generally.
Future work should seek to understand precisely when ATAs are or are not important
for coexistence.

%Para continuing to try to build intuition about how ATAs work
We return to the task, begun in the Introduction, of building intuition for how
ATA effects work. 
Just as for storage effects [\cite{Chesson_1994}, Box 2 of \cite{StumpVasseur2023}], 
temporal niche differentiation is necessary for ATA
effects (SI section \ref{SI-SIsect:temporal_niche_diff}). 
If temporal niche differentiation occurs
but is partial, as will be typical, $E_i$ and $C_{i \bs i}$ for a 
species $i$ will be positively but imperfectly associated, 
and the association may be asymmetric. 
Storage events occur for species $i$ when the positive effects of $E_i$ overwhelm the negative 
effects of $C_{i \bs i}$, producing large values of $r_i(E_i,C_{i \bs i})$
which are not completely offset by population declines at other times. 
For instance, for the lottery model,
$r_i(E_i,C_{i \bs i})=\log(1-\delta+E_i/C_{i \bs i})$ can get large whenever
$C_{i \bs i}$ is a small fraction of $E_i$, but can never get smaller
than $\log(1-\delta)$. 
The potential for storage events can differ when $E_i$ and $C_{i \bs i}$ are both
above their median values compared to when they are both below median. For instance,
the ratio $E_i/C_{i \bs i}$, and therefore the growth rate 
$\log(1-\delta+E_i/C_{i \bs i})$, can both get very big, corresponding to
rapid growth, if $C_{i \bs i}$ is close to $0$, even if $E_i$ is fairly small.
In contrast, if $E_i$ and $C_{i \bs i}$ are both in the right portions of their distributions, 
so that $C_{i \bs i}$ is bounded away from $0$, there may be much sharper limits on the 
ratio $E_i/C_{i \bs i}$ and the growth rate.
ATA effects quantify the extent to which asymmetries of association between $E$ and 
$C$ are aligned with varying potential for storage across the $E,C$ distribution. 
ATA effects help quantify different degrees of temporal niche differentiation 
occurring under different sets of environmental conditions, possibly via
different mechanisms.
Further details and examples of this intuition 
are in SI section \ref{SI-SIsect:intuition_examples} and SI 
Figs \ref{SI-fig:beta_fecundidities_explanatory}-\ref{SI-fig:lognormal_fecundidities_explanatory}, 
where comparisons are also made between our two lottery models.
